Visual Diagnostics of a Model Explainer -- Tools for the Assessment of LIME Explanations

### 2020-08-27 paper.Rnw
edits through the discussion

LIME \sout{is implemented using the R package \emph{limeaid} to apply LIME} \kge{is applied} \sout{multiple times} to all test set observations using different tuning parameters: 12 sampling methods (2-6 equally spaced bins, 2-6 quantile bins, kernel density estimation, and normal approximation) and 3 Gower exponents (0.5, 1, and 10). Thus, a total of $12\times 3=36$ different applications of LIME were performed. We specify that each LIME explanation return 3 features and feature selection is performed \kge{by selecting the features with the highest weights in a ridge regression model} \sout{using the highest weights method} (default option). \kgc{XXX THIS IS WHERE I LEFT OFF XXX}

@@

To get an overview of the LIME explanations from the 36 applications, we consider a feature heatmap (\autoref{fig:bullet-feature-heatmap}). In addition to facets for simulation method and order of feature selected by LIME, this plot includes a vertical facet for Gower power and a horizontal facet for whether the observation is a known match or non-match. This plot highlights several key features of the LIME explanations from the bullet matching dataset.

@@

@
\caption{Feature heatmap of the 36 LIME applications to the bullet comparison data test set. In addition to faceting the results by simulation method and LIME feature selection order, facets for the Gower power and whether the observation is a match or non-match are included. The vertical stripes of features selected indicate a dependence between the LIME explanations and tuning parameters.}
\label{fig:bullet-feature-heatmap}

@@

# Create a metric comparison plot
plot_metrics(bullet_explain %>% 
               mutate(label = as.factor(label)), 
             add_lines = TRUE) +
  theme_bw(base_family = "Times") +
  theme(strip.background = element_rect(color = "white", fill = "white"))
@
\caption{Plot of LIME assessment metrics for the applications of LIME on the bullet comparison data test set. The density simulation methods perform well for all metrics. The bin based metrics often do not agree in terms of performance across metrics.}
\label{fig:bullet-assess-metric}
\end{figure*}

@@

Figure \ref{fig:bullet-eois} include the explanation plots (from \emph{lime}) and explanation scatterplots for cases M and NM. Plots for 3 equal bins and 3 quantile bins are included. Plots for the kernel density simulation method are included in Figure \ref{fig:bullet-expl-scat-density}. The scatterplots provide several insights into the LIME explanations. Here we focus on the explanation for case NM with 3 quantile bins, but similar statements can be made about the other plots.

The explanation plot from \emph{lime} for case NM with 3 quantile bins shows that an observed CCF value greater than 0.320 supports a random forest prediction in favor of a match, and the scatter plots show that many of the simulated values with CCF greater then 0.320 have random forest predictions greater than 0.5. Additionally, the LIME explanation indicates that a value of matches less than or equal to 1.66 contradicts a prediction in favor of a match, and the scatter plots show that almost all simulated values with a value of matches less than or equal to 1.66 have random forest predictions close to 0. While the LIME explanation makes sense based on the observations made from the explanation scatter plot, the plot also indicates that LIME falls short of providing a good explanation of the random forest prediction for case NM. Based on the scatter plot of CCF versus matches, a better explanation would be that because case NM has a value of CCF less than 0.8 and a value of matches less than 7, the random forest provides a prediction supporting a non-match. The relationship between CCF and rough correlation does not provide much evidence to support the random forest prediction one way or the other, but there is a pentagon shaped region at the bottom of the scatter plot of matches versus rough correlation with mostly predictions close to 0 that case NM falls and supports the random forest prediction of a non-match.

@@

# eoi1_kd <-
#   plot_explain_scatter(bullet_explain_perms[13:15, ], alpha = 0.9, weights = TRUE) +
#   scale_color_gradient2(
#     low = "grey50",
#     high = "darkorange",
#     midpoint = 0.5,
#     limits = c(0, 1)
#   ) +
#   scale_fill_gradient2(
#     low = "grey50",
#     high = "darkorange",
#     midpoint = 0.5,
#     limits = c(0, 1)
#   ) +
#   theme_bw(base_family = "Times", base_size = 8) +
#   theme(
#     strip.placement = "outside",
#     strip.background = element_rect(color = "white",
#                                     fill = "white")
#   )
# 
# eoi2_kd <-
#   plot_explain_scatter(bullet_explain_perms[16:18, ], alpha = 0.9, weights = TRUE) +
#   scale_color_gradient2(
#     low = "grey50",
#     high = "darkorange",
#     midpoint = 0.5,
#     limits = c(0, 1)
#   ) +
#   scale_fill_gradient2(
#     low = "grey45",
#     high = "darkorange",
#     midpoint = 0.5,
#     limits = c(0, 1)
#   ) +
#   theme_bw(base_family = "Times", base_size = 8) +
#   theme(
#     strip.placement = "outside",
#     strip.background = element_rect(color = "white",
#                                     fill = "white")
#   )

pf1_3qb <- plot_features(bullet_explain_perms[1:3,]) +
  scale_fill_manual(values = c("darkorange", "grey50")) +
  theme(text = element_text(family = "Times", size = 10))
pf2_3qb <- plot_features(bullet_explain_perms[4:6,]) +
  scale_fill_manual(values = c("darkorange", "grey50")) + 
  theme(text = element_text(family = "Times", size = 10))
pf1_3eb <- plot_features(bullet_explain_perms[7:9,]) +
  scale_fill_manual(values = c("darkorange", "grey50")) +
  theme(text = element_text(family = "Times", size = 10))
pf2_3eb <- plot_features(bullet_explain_perms[10:12,]) +
  scale_fill_manual(values = c("darkorange", "grey50")) +
  theme(text = element_text(family = "Times", size = 10))
# pf1_kd <- plot_features(bullet_explain_perms[13:15,]) +
#   scale_fill_manual(values = c("darkorange", "grey50"))
# pf2_kd <- plot_features(bullet_explain_perms[16:18,]) +
#   scale_fill_manual(values = c("darkorange", "grey50"))
@

\begin{figure*}[!thp]
<<bullet-eois, out.width = '\\textwidth', fig.width = 12, fig.height = 17, fig.align = "center", out.width="90%", message = FALSE>>=
# Join the plots
plot_grid(
  pf2_3qb,
  pf2_3eb,
  eoi2_3qb,
  eoi2_3eb,
  pf1_3qb,
  pf1_3eb,
  eoi1_3qb,
  eoi1_3eb,
  nrow = 4,
  rel_heights = c(0.2, 0.3, 0.2, 0.3)
)
@
\caption{Plots of LIME explanations and explanation scatterplots for one case M and NM in the bullet test data. The first column contains plots associated with 3 quantile bins and the second with 3 equally spaced bins. The plots provide insights into the LIME explanations and allow for the assessment of the explanation quality.}
\label{fig:bullet-eois}
\end{figure*}

Without applying LIME with multiple tuning parameters to the bullet test data or viewing diagnostic plots of the LIME explanations, it may be very possible to formulate reasons why the LIME explanations make sense. However, the sequence of plots in this section (Figures \ref{fig:bullet-feature-heatmap}, \ref{fig:bullet-assess-metric}, and \ref{fig:bullet-eois}) suggest that we should be cautious to trust any of these LIME explanations. It appears that either LIME needs to be further tuned to provide trustworthy and good explanations, or a different approach may provide better insight.

@@

%\hh{The discussion should not have subsection ... but we can keep them for now for structural purposes.} \kg{Okay. I took them out.}

This paper highlights that while an explainer model is meant to provide clarity, it actually adds another layer of complexity to predictive models by requiring yet another model that needs to be assessed. Without an assessment of the explainer model, LIME is a black-box procedure of its own requiring blind trust in the explainer model. %\hh{XXX what is the implication? bring it to a punch: asking the user to trust into explainer...}
We suggest the use of visual diagnostics to counteract the black-box nature of LIME and provide three diagnostic plots.

The visualizations are intended to provide insight on how LIME works, assess the ability of the explainer model to capture the complex predictive model, and compare LIME explanations produced by different tuning parameters. While the visualizations accomplish these tasks, they also expose examples of the failings of LIME. To address the discovered failings of LIME, we will reconsider each of the claims about the performance of LIME made by \citet{ribeiro:2016} in light of the insights gained from the diagnostic visualizations.

As previously discussed, the \textbf{interpretability} of the LIME explanations can be controlled by the complexity of the explainer model. For example, the number of bins selected for simulation can control the interpretability of the explanations. If too many bins are selected, the bin range that is reported in the LIME explanation will be too small to be meaningful in the context of the feature. An appropriate choice of the number of bins will keep the bin range meaningful. Thus, the claim of interpretability does not need to be assessed using the visualizations. However, diagnostic visualizations do present a different perspective on the meaning of interpretability.

Even though an explanation will be interpretable as long as the complexity of the explainer model is appropriately chosen, a lack of understanding of the process used to create the explanation could lead to an incorrect interpretation of the explanation. For example, the visualization of a LIME explanation available from the \emph{lime} R package \citep{pedersen:2020} (shown in Figures \ref{fig:sine-exp-scatter} and \ref{fig:bullet-eois}) is a major simplification of the explainer model, which could lead to under-interpreted or misinterpreted LIME explanation. Supplementing \citet{pedersen:2020}'s compact visualization of the explanation with an explanation scatterplot that shows a more detailed visualization of the explainer model (such as Figures \ref{fig:sine-exp-scatter} and \ref{fig:bullet-eois}) promotes a full interpretation of the explanation.

Even with an explainer model that is interpreted correctly, the interpretation is worthless if the explainer model is not \textbf{faithful} to the complex model. This claim can be assessed using the diagnostic plots suggested in this paper. Many of the visualizations in this paper highlight problems with the faithfulness of the explainer models. The explanation scatterplots allow for a comparison of the explainer model to the complex model. The examples in this paper show cases where the explainer model bins do not accurately capture the classification boundaries near the predictions of interest of the random forest\kge{s \sout{models}} by oversimplifying the model (Figures \ref{fig:sine-exp-scatter} and \ref{fig:bullet-eois}). Using less bins would clearly not help improve the faithfulness of the explainer model in these examples, and while an increase in bins would lead to a finer resolution of the random forest classification boundaries, interpretability of the explainer model would quickly be lost. Perhaps this could be improved by allowing the bin creation to account for the relationships between the features and response variable or a different number of bins for each feature.

In addition to assessing faithfulness by visually comparing the complex and explainer model, we propose a visual comparison of two faithfulness metrics (MSEE and average fidelity). The examples of faithfulness metric comparisons in this paper (Figures \ref{fig:sine-assess-metric} and \ref{fig:bullet-assess-metric}) both produced confusing results. The density based simulation methods resulted in the best or close to the best performance even though the feature heatmaps (Figures \ref{fig:sine-feat-heat} and \ref{fig:bullet-feature-heatmap}) showed that the density simulation methods produced global explanations with no variation in features selected by LIME. For the bin based simulation methods, the two metrics often did not agree or contradicted one another, which makes it difficult to decide on a recommendation of a set of tuning parameters that produces the explanations with the most faithful explainer model.

@@

The final claim, \textbf{localness}, is addressed by the feature heatmap and metric comparison plot. As stated, the feature heatmap revealed that the density simulation methods in the examples of this paper resulted in global explanations where the same features were repeatedly chosen across all (or almost all) observations in the set of explanations. This finding agrees with that of \citep{laugel:2018} who found LIME produced global explanations with the normal approximation simulation method. For the bin based simulation methods in the bullet data example, the feature heatmap showed that the features chosen for the explanations varied between the two classification categories (match versus non-match). This is an interesting finding that suggests that different features can play a role in the predictions of observations in different response categories, but the patterns of features selected by LIME within a classification category do not vary. Again, this is a suggestion of global explanations. Furthermore, while the metric comparison plot in the bullet example did not provide agreement between metrics on a best bin based method, all metrics agree that a Gower power of 0.5 for computing the model weights associated with distance of a simulated data point from the prediction of interest is best. This suggests that a less local explanation provided a better explanation of the performance of the random forest.

Some of the visualizations in the paper generalize easily to any application of LIME such as the feature heatmap and metric plot. Other plots such as the visualizations of the LIME procedure would require extensions such as the use of scatterplot matrices to compare explanations with more than two features. The addition of interactivity to the diagnostic plots would provide additional enhancement of the assessment process. For example, a diagnostic plot that provides a summary of multiple LIME explanations, such as the feature heatmap, could be displayed and clicked on to reveal more detailed figures associated with individual predictions of interest, such as plots of the simulated data and explainer model.

The largest limitation to the diagnostic visualizations is the dimensionality of the data shown, both in the number of dimensions or features as well as the number of observations. Fortunately, in the situation of LIME, both of these aspects are rather well controlled: LIME relies heavily on simulations to generate data scenarios that are close to the data observed but exhibits variability. Effects from overplotting should be relatively mild, because output from simulations is shown, which is expected to be (relatively) continuous \hh{such that overplotting only occurs for points with (relatively) similar values.} \kgc{What do you mean by continuous here?}. \hh{XXX Does that extra explanation help?} \kgc{Yes. I see what you mean now.} In that respect, the diagnostics shown for the \data\ and in the bullet example are representative of what is expected. But in cases where overplotting does become problematic, the user could either simply reduce the size of the simulations or use some well-studied binning techniques in the visualizations, as discussed for example in \citet{carr:1987} or \citet{unwin:2006}.

While it would be ideal if LIME could be used as a method to provide easily understandable explanations for black-box models as \citep{ribeiro:2016} claim, that dream is not yet a reality. The examples using diagnostic plots to assess LIME in this paper show frequent issues with LIME. We hope that our plots provide motivation to assess LIME explanations, to not blindly use the default settings (even if it is not clear which tuning parameters to use), and to encourage work on improving LIME, so that it can be a lime and not a lemon.

\section*{Acknowledgments}
HH was partially funded by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreement 70NANB20H019 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, Duke University, University of California Irvine, University of Virginia, West Virginia University, University of Pennsylvania, Swarthmore College and University of Nebraska, Lincoln.

@@

\hh{A taxonomy of author contributions is suggested by Elsevier - I put the link in an email
}

{\bf Katherine Goode, Heike Hofmann}: Conceptualization, Methodology, Investigation, Writing- Reviewing and Editing.
{\bf Katherine Goode}: Writing Initial Draft, Software.

@@

The following supporting information is available as part of the online article:

@@

% \textsc{\textbf{Ideas for describing proposed plots}}
% 
% \begin{itemize}
% \item start at the high level explanation (this is what we want to do and this is how we do it)
% \item go over a good and bad example
% \item include schematics of the plot first (good and bad) and then include the sine mini example version
% \item explain the expectation of the plot
% \end{itemize}
% 
% \textsc{\textbf{Comments for Editing}}
% 
% \begin{itemize}
% \item \hh{Generally: avoid 'can be'. Replace by 'is'.  }
% \item \hh{References like 'they', 'them' ... replace them by repeating the noun you are referring to to avoid any kind of ambiguity}
% \item \kgc{words to go back and make sure I am not using too often: ability, produce, understand}
% \item \kgc{decide on features or predictor variables}
% \item \hh{it's tuning parameters, not tunning} \kgc{Oops! I get that one wrong all of the time}
% \item \kgc{Adjust text sizes in plots so they all match}
% \item \kgc{Color scheme - comment from Heike:} \hh{I like the red-blue scheme for binary values, I also like the color scheme in the features heatmap. Generally papers and color don't go well together anyways, but we will solve that problem when we have to ... Most likely we will have to just pay the extra color fee}
% \item \kgc{Check that I am consistently referring to the perturbing as data simulation}
% \item \kgc{Clean up notation}
% \item \kgc{Referring to the linearity assumption:} \hh{Be a bit careful in the phrasing here - we can always approximate any function using a linear form (think Taylor expansion). The question is how big the error is.}
% \item \kgc{Make sure plots in the same figure are the same size}
% \item \kgc{Make sure I am distinguishing between RF predictions and probabilities correctly}
% \item \kgc{Use prediction of interest not case of interest}
% \item \kgc{Make sure I always treat the word 'data' as plural}
% \item \kgc{dataset or data set?}
% \item \kgc{scatterplot or scatter plot (check other plot names as well)}
% \item \kgc{choose whether to use the term feature or variable and make sure it is consistent throughout}
% \item \kgc{just include the details about LIME that are necessary to understand the plots and remove the rest to the appendix}
% \item \hh{avoid the use of implicit references like `this' and `that'. Make the reference explicit by adding what you are referring to, such as `this method', `that procedure'}
% \item \hh{make sure to use exact phrases. things like `LIME was developed in 2016' is factually questionable - but we know LIME was introduced in 2016. Also `the authors were interested in ...' implies that you know their intent. It's better to not guess any intent but stick to what we know.}
% \item \hh{`our` diagnostics ... you are using `our` quite often. It is good to claim work, but it is not good to claim things that we want others to use, because that is by definition excluding everybody but you. It's better to re-phrase with a focus on what you did. We introduce, we suggest, etc.}
% \item \kgc{check that I am always using present tense}
% \item \kgc{check that I am consistent in my use of the term black-box model}
% \end{itemize}
\newpage
\appendix

>>>>

LIME \sout{is implemented using the R package \emph{limeaid} to apply LIME} \kge{is applied} \sout{multiple times} to all test set observations using different tuning parameters: 12 sampling methods (2-6 equally spaced bins, 2-6 quantile bins, kernel density estimation, and normal approximation) and 3 Gower exponents (0.5, 1, and 10). Thus, a total of $12\times 3=36$ different applications of LIME were performed. We specify that each LIME explanation return 3 features and feature selection is performed \kge{by selecting the features with the highest weights in a ridge regression model} \sout{using the highest weights method} (default option).

@@

To get an overview of the LIME explanations from the 36 applications, we consider a feature heatmap (\autoref{fig:bullet-feature-heatmap}). In addition to facets for simulation method and \kge{LIME feature importance} \sout{order of feature selected by LIME}, this plot includes a vertical facet for Gower power and a horizontal facet for whether the observation is a known match or non-match. This plot highlights several key features of the LIME explanations from the bullet matching dataset.

@@

@
\caption{Feature heatmap of \sout{the} 36 LIME applications to the bullet comparison data test set. In addition to faceting the results by simulation method and LIME feature selection order, facets for the Gower power and whether the observation is a match or non-match are included. The vertical stripes of features selected indicate a dependence between the LIME explanations and tuning parameters.}
\label{fig:bullet-feature-heatmap}

@@

# Create a metric comparison plot
plot_metrics(bullet_explain %>% 
               mutate(label = as.factor(label)), 
             add_lines = TRUE) +
  theme_bw(base_family = "Times") +
  theme(strip.background = element_rect(color = "white", fill = "white"))
@
\caption{\kge{Assessment metric plot} \sout{Plot of LIME assessment metrics} for the applications of LIME on the bullet comparison data test set. The density simulation methods perform well for all metrics. The bin based metrics often do not agree in terms of performance across metrics.}
\label{fig:bullet-assess-metric}
\end{figure*}

@@

Figure \ref{fig:bullet-eois} include\kge{s} the \kge{visual representation plots of explanations} \sout{explanation plots} (from \emph{lime}) and explanation scatterplots for cases M and NM. Plots for 3 equal bins and 3 quantile bins are included. Plots for the kernel density simulation method are included in \kge{Appendix \ref{exp-scatter_plus} (}Figure \ref{fig:bullet-expl-scat-density}\kge{)}. The scatterplots provide several insights into the LIME explanations. Here we focus on the explanation for case NM with 3 quantile bins, but similar statements can be made about the other plots. \kgc{I've been thinking about the examples I chose to include, and I have a couple of questions. First, is it okay to include all of these plots when I am only discussing one? Second, I'm not sure why I chose to list 3 quantile bins as a "good" tuning parameter value anymore based on the assessment metric plot. I don't know that I agree with that anymore. However, it does provide a nice example for where the explainer model goes wrong. I'm wondering if it would be better to just focus on the 3 equal bins plots for cases M and NM. They are not quite a dramatically bad as 3 quantile bins, but a case could be make for not really capturing the random forest prediction regions.}

The explanation plot from \emph{lime} for case NM with 3 quantile bins shows that an observed CCF value greater than 0.320 supports a random forest prediction in favor of a match, and the scatter plots show that many of the simulated values with CCF greater then 0.320 have random forest predictions greater than 0.5. Additionally, the LIME explanation indicates that a value of matches less than or equal to 1.66 contradicts a prediction in favor of a match, and the scatter plots show that almost all simulated values with a value of matches less than or equal to 1.66 have random forest predictions close to 0. While the LIME explanation makes sense based on the observations made from the explanation scatter plot, the plot also indicates that LIME falls short of providing a good explanation of the random forest prediction for case NM. Based on the scatter plot of CCF versus matches, a better explanation would be that because case NM has a value of CCF less than 0.8 and a value of matches less than 7, the random forest provides a prediction supporting a non-match. The relationship between CCF and rough correlation does not provide much evidence to support the random forest prediction one way or the other, but there is a pentagon shaped region at the bottom of the scatter plot of matches versus rough correlation with mostly predictions close to 0 that case NM falls \kge{in} and supports the random forest prediction of a non-match.

@@

pf1_3qb <- plot_features(bullet_explain_perms[1:3,]) +
  scale_fill_manual(values = c("darkorange", "grey50")) +
  theme(text = element_text(family = "Times", size = 10))
pf2_3qb <- plot_features(bullet_explain_perms[4:6,]) +
  scale_fill_manual(values = c("darkorange", "grey50")) + 
  theme(text = element_text(family = "Times", size = 10))
pf1_3eb <- plot_features(bullet_explain_perms[7:9,]) +
  scale_fill_manual(values = c("darkorange", "grey50")) +
  theme(text = element_text(family = "Times", size = 10))
pf2_3eb <- plot_features(bullet_explain_perms[10:12,]) +
  scale_fill_manual(values = c("darkorange", "grey50")) +
  theme(text = element_text(family = "Times", size = 10))
@

\begin{figure*}[!thp]
<<bullet-eois, out.width = '\\textwidth', fig.width = 12, fig.height = 17, fig.align = "center", out.width="90%", message = FALSE>>=
# Join the plots
plot_grid(
  pf2_3qb,
  pf2_3eb,
  eoi2_3qb,
  eoi2_3eb,
  pf1_3qb,
  pf1_3eb,
  eoi1_3qb,
  eoi1_3eb,
  nrow = 4,
  rel_heights = c(0.2, 0.3, 0.2, 0.3)
)
@
\caption{Plots of LIME explanations \kge{(first and third rows)} and explanation scatterplots \kge{(second and fourth rows)} for \sout{one} case\kge{s} M and NM in the bullet test data \kge{for two tuning parameter values: 3 quantile bins (first column) and 3 equal bins (second column)}. \sout{The first column contains plots associated with 3 quantile bins and the second with 3 equally spaced bins.} The plots provide insights into the LIME explanations and allow for the assessment of the explanation quality.}
\label{fig:bullet-eois}
\end{figure*}

Without applying LIME with multiple tuning parameter\sout{s} \kge{values} to the bullet test data or viewing diagnostic plots of the LIME explanations, it may be very possible to formulate reasons why the LIME explanations make sense. However, the sequence of plots in this section (Figures \ref{fig:bullet-feature-heatmap}, \ref{fig:bullet-assess-metric}, and \ref{fig:bullet-eois}) suggest that we should be cautious to trust any of these LIME explanations. It appears that either LIME needs to be further tuned to provide trustworthy and good explanations, or a different approach may provide better insight.

@@

This paper highlights that while an explainer model is meant to provide clarity, it actually adds another layer of complexity to predictive models by requiring yet another model that needs to be assessed. Without an assessment of the explainer model, LIME is a black-box procedure of its own requiring blind trust in the explainer model. We suggest the use of visual diagnostics to counteract the black-box nature of LIME and provide three diagnostic plots.

The visualizations are intended to provide insight on how LIME works, assess the ability of the explainer model to capture the complex predictive model, and compare LIME explanations produced by different tuning parameter\sout{s} \kge{values}. While the visualizations accomplish these tasks, they also expose examples of the failings of LIME. To address the discovered failings of LIME, we \sout{will} reconsider each of the claims about the performance of LIME made by \citet{ribeiro:2016} in light of the insights gained from the diagnostic visualizations.

As previously discussed, the \textbf{interpretability} of the LIME explanations \kge{is controllable} \sout{can be controlled} by the complexity of the explainer model. For example, the number of bins selected for simulation can control the interpretability of the explanations. If too many bins are selected, the bin range that is reported in the LIME explanation will be too small to be meaningful in the context of the feature. An appropriate choice of the number of bins will keep the bin range meaningful. Thus, the claim of interpretability does not need to be assessed using the visualizations. However, diagnostic visualizations do present a different perspective on the meaning of interpretability.

Even though an explanation will be interpretable as long as the complexity of the explainer model is appropriately chosen, a lack of understanding of the process used to create the explanation could lead to an incorrect interpretation of the explanation. For example, the visualization of a LIME explanation available from the \emph{lime} R package \citep{pedersen:2020} (shown in Figures \ref{fig:sine-exp-scatter} and \ref{fig:bullet-eois}) is a major simplification of the explainer model, which could lead to under-interpreted or misinterpreted LIME explanation. Supplementing \citet{pedersen:2020}'s compact visualization of the explanation with an explanation scatterplot that shows a more detailed visualization of the explainer model (such as Figures \ref{fig:sine-exp-scatter} and \ref{fig:bullet-eois}) promotes a \kge{more complete understanding} \sout{full interpretation} of the explanation.

Even with an explainer model that is interpreted correctly, the interpretation is worthless if the explainer model is not \textbf{faithful} to the complex model. This claim can be assessed using the diagnostic plots suggested in this paper. Many of the visualizations in this paper highlight problems with the faithfulness of the explainer models. The explanation scatterplots allow for a comparison of the explainer model to the complex model. The examples in this paper show cases where the explainer model bins do not accurately capture the \kge{regions with similar random forest probabilities that contain the prediction of interest} \sout{classification boundaries near the predictions of interest of the random forest\kge{s \sout{models}} by} \kge{and} oversimplify\sout{ing} the model (Figures \ref{fig:sine-exp-scatter} and \ref{fig:bullet-eois}). Using less bins would clearly not help improve the faithfulness of the explainer model in these examples, and while an increase in bins would lead to a finer resolution of the random forest classification boundaries, interpretability of the explainer model would quickly be lost. Perhaps this could be improved by allowing the bin creation to account for the relationships between the features and response variable or a different number of bins for each feature.

In addition to assessing faithfulness by visually comparing the complex and explainer model, we propose a visual comparison of two faithfulness metrics (MSEE and average fidelity). The examples of faithfulness metric comparisons in this paper (Figures \ref{fig:sine-assess-metric} and \ref{fig:bullet-assess-metric}) both produced confusing results. The density based simulation methods result\sout{ed} in the best or close to the best performance even though the feature heatmaps (Figures \ref{fig:sine-feat-heat} and \ref{fig:bullet-feature-heatmap}) show\sout{ed} that the density simulation methods produce\sout{d} global explanations with \kge{minimal} \sout{no} variation in features selected by LIME. For the bin based simulation methods, the two metrics often \kge{do} \sout{did} not agree or contradict\sout{ed} one another, which makes it difficult to decide on a recommendation of a set of tuning parameters that produces the explanations with the most faithful explainer model.

@@

The final claim, \textbf{localness}, is addressed by the feature heatmap and metric comparison plot\kge{s}. As stated, the feature heatmap reveal\kge{s}\sout{ed} that the density simulation methods in the examples of this paper result\kge{s}\sout{ed} in global explanations where the same features \kge{are} \sout{were} repeatedly chosen across all (or almost all) observations in the set of explanations. This finding agrees with that of \citep{laugel:2018} who found LIME produced global explanations with the normal approximation simulation method. For the bin based simulation methods in the bullet data example, the feature heatmap showed that the features chosen for the explanations varied between the two classification categories (match versus non-match). This is an interesting finding that suggests that different features can play a role in the predictions of observations in different response categories\kge{.} \sout{, but the patterns of features selected by LIME within a classification category do not vary. Again, this is a suggestion of global explanations.} \kgc{I'm not sure what I was trying to say there.} Furthermore, while the metric comparison plot in the bullet example did not provide agreement between metrics on a best bin based method, all metrics agree that a Gower power of 0.5 for computing the model weights associated with distance of a simulated data point from the prediction of interest is best. This suggests that a less local explanation provided a better explanation of the performance of the random forest.

Some of the visualizations in the paper generalize easily to any application of LIME such as the feature heatmap and metric plot. Other plots such as the visualizations of the LIME procedure would require extensions such as the use of scatterplot matrices to compare explanations with more than two features. The addition of interactivity to the diagnostic plots would provide additional enhancement of the assessment process. For example, a diagnostic plot that provides a summary of multiple LIME explanations, such as the feature heatmap, could be displayed and clicked on to reveal more detailed figures associated with individual predictions of interest, such as \kge{an explanation scatterplot.} \sout{plots of the simulated data and explainer model.}

The largest limitation to the diagnostic visualizations is the dimensionality of the data shown, both in the number of dimensions or features as well as the number of observations. Fortunately, in the situation of LIME, both of these aspects are rather well controlled: LIME relies heavily on simulations to generate data scenarios that are close to the data observed but exhibits variability. Effects from overplotting should be relatively mild, because output from simulations is shown, which is expected to be (relatively) continuous \hh{such that overplotting only occurs for points with (relatively) similar values.} \kgc{What do you mean by continuous here?}. \hh{XXX Does that extra explanation help?} \kgc{Yes. I see what you mean now.} In that respect, the diagnostics shown for the \data \ and \sout{in} the bullet example are representative of what is expected. But in cases where overplotting does become problematic, the user could either simply reduce the size of the simulations or use some well-studied binning techniques in the visualizations, as discussed for example in \citet{carr:1987} or \citet{unwin:2006}.

While it would be ideal if LIME could be used as a method to provide easily understandable explanations for black-box models as \citet{ribeiro:2016} claim, that dream is not yet a reality. The examples using diagnostic plots to assess LIME in this paper show frequent issues with LIME. We hope that our plots provide motivation to assess LIME explanations, to not blindly use the default settings (even if it is not clear \kge{how to tune the parameters} \sout{which tuning parameters to use}), and to encourage work on improving LIME, so that it can be a lime and not a lemon.

\section*{Acknowledgments}

\kge{Should this go under the financial disclosure section? Also, what is the financial disclosure section meant to contain? I couldn't find it described on the ASA data science journal author guidelines page. Should we remove it?} HH was partially funded by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreement 70NANB20H019 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, Duke University, University of California Irvine, University of Virginia, West Virginia University, University of Pennsylvania, Swarthmore College and University of Nebraska, Lincoln.

@@

\hh{A taxonomy of author contributions is suggested by Elsevier - I put the link in an email}

{\bf Katherine Goode, Heike Hofmann}: Conceptualization, Methodology, Investigation\kge{, Visualizations}, Writing- Reviewing and Editing.
{\bf Katherine Goode}: Writing Initial Draft, Software.

@@

The following supporting information is available as part of the online article: \kgc{What should we list here?}

@@

% \textsc{\textbf{Ideas for describing proposed plots}}
% 
% \begin{itemize}
% \item start at the high level explanation (this is what we want to do and this is how we do it)
% \item go over a good and bad example
% \item include schematics of the plot first (good and bad) and then include the sine mini example version
% \item explain the expectation of the plot
% \end{itemize}
% 
% \textsc{\textbf{Comments for Editing}}
% 
% \begin{itemize}
% \item \hh{Generally: avoid 'can be'. Replace by 'is'.  }
% \item \hh{References like 'they', 'them' ... replace them by repeating the noun you are referring to to avoid any kind of ambiguity}
% \item \kgc{words to go back and make sure I am not using too often: ability, produce, understand}
% \item \kgc{decide on features or predictor variables}
% \item \hh{it's tuning parameters, not tunning} \kgc{Oops! I get that one wrong all of the time}
% \item \kgc{Adjust text sizes in plots so they all match}
% \item \kgc{Color scheme - comment from Heike:} \hh{I like the red-blue scheme for binary values, I also like the color scheme in the features heatmap. Generally papers and color don't go well together anyways, but we will solve that problem when we have to ... Most likely we will have to just pay the extra color fee}
% \item \kgc{Check that I am consistently referring to the perturbing as data simulation}
% \item \kgc{Clean up notation}
% \item \kgc{Referring to the linearity assumption:} \hh{Be a bit careful in the phrasing here - we can always approximate any function using a linear form (think Taylor expansion). The question is how big the error is.}
% \item \kgc{Make sure plots in the same figure are the same size}
% \item \kgc{Make sure I am distinguishing between RF predictions and probabilities correctly}
% \item \kgc{Use prediction of interest not case of interest}
% \item \kgc{Make sure I always treat the word 'data' as plural}
% \item \kgc{dataset or data set?}
% \item \kgc{scatterplot or scatter plot (check other plot names as well)}
% \item \kgc{choose whether to use the term feature or variable and make sure it is consistent throughout}
% \item \kgc{just include the details about LIME that are necessary to understand the plots and remove the rest to the appendix}
% \item \hh{avoid the use of implicit references like `this' and `that'. Make the reference explicit by adding what you are referring to, such as `this method', `that procedure'}
% \item \hh{make sure to use exact phrases. things like `LIME was developed in 2016' is factually questionable - but we know LIME was introduced in 2016. Also `the authors were interested in ...' implies that you know their intent. It's better to not guess any intent but stick to what we know.}
% \item \hh{`our` diagnostics ... you are using `our` quite often. It is good to claim work, but it is not good to claim things that we want others to use, because that is by definition excluding everybody but you. It's better to re-phrase with a focus on what you did. We introduce, we suggest, etc.}
% \item \kgc{check that I am always using present tense}
% \item \kgc{check that I am consistent in my use of the term black-box model}
% \end{itemize}

\newpage

\appendix

\kgc{XXX This is where I left off XXX}
